---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Riccardo Fusaroli"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Analysing Heart Rate and Respiration data

The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from last year (Study1) and from your class (Study2). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: last year it was self-paced joint reading; this year it was the tv-series conversation. So you might want to exclude the self-paced reading (but, up to you!)

## Step by step suggestions to solve the assignment


### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
```{r}
pair6_con1=read.csv("cleandata/Study2_G6_T1_TurnTaking.csv")
pair6_con2=read.csv("cleandata/Study2_G6_T2_Conversation.csv")
pair6_con3=read.csv("cleandata/Study2_G6_T3_Synchronous.csv")
```


- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)
  
```{r}

#remove outliers

removeOuts <- function(ts,threshold){
  ts[ts > (mean(ts,na.rm=T) +
          (threshold*sd(ts,na.rm=T))) |
     ts < (mean(ts,na.rm=T) -
          (threshold*sd(ts,na.rm=T)))] =
  mean(ts,na.rm=T)   
  return(ts)
  }

threshold=1.5
pair6_con1$HR1=removeOuts(pair6_con1$HR1,threshold)
pair6_con2$HR1=removeOuts(pair6_con2$HR1,threshold) 
pair6_con3$HR1=removeOuts(pair6_con3$HR1,threshold)

pair6_con1$HR2=removeOuts(pair6_con1$HR2,threshold)
pair6_con2$HR2=removeOuts(pair6_con2$HR2,threshold) 
pair6_con3$HR2=removeOuts(pair6_con3$HR2,threshold)

pair6_con1$Resp1=removeOuts(pair6_con1$Resp1,threshold)
pair6_con2$Resp1=removeOuts(pair6_con2$Resp1,threshold) 
pair6_con3$Resp1=removeOuts(pair6_con3$Resp1,threshold)

pair6_con1$Resp2=removeOuts(pair6_con1$Resp2,threshold)
pair6_con2$Resp2=removeOuts(pair6_con2$Resp2,threshold) 
pair6_con3$Resp2=removeOuts(pair6_con3$Resp2,threshold)


#downsampling

library(groupdata2)
library(tidyverse)
d_pair6_con1 = pair6_con1 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(
    time= mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

d_pair6_con2 = pair6_con2 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(
    time= mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))

d_pair6_con3 = pair6_con3 %>%    
  group(n= 100, method= 'greedy') %>%    
  dplyr::summarise(
    time= mean(time,na.rm=T),
    HR1 = mean(HR1,na.rm=T),
    HR2 = mean(HR2,na.rm=T),
    Resp1 = mean(Resp1,na.rm=T),
    Resp2 = mean(Resp2,na.rm=T))


#scaling

d_pair6_con1$Resp1S=scale(d_pair6_con1$Resp1)
d_pair6_con1$Resp2S=scale(d_pair6_con1$Resp2)
d_pair6_con1$HR1S=scale(d_pair6_con1$HR1) 
d_pair6_con1$HR2S=scale(d_pair6_con1$HR2)

d_pair6_con2$Resp1S=scale(d_pair6_con2$Resp1)
d_pair6_con2$Resp2S=scale(d_pair6_con2$Resp2)
d_pair6_con2$HR1S=scale(d_pair6_con2$HR1) 
d_pair6_con2$HR2S=scale(d_pair6_con2$HR2)

d_pair6_con3$Resp1S=scale(d_pair6_con3$Resp1)
d_pair6_con3$Resp2S=scale(d_pair6_con3$Resp2)
d_pair6_con3$HR1S=scale(d_pair6_con3$HR1) 
d_pair6_con3$HR2S=scale(d_pair6_con3$HR2)

#plot
library(ggplot2)

#heartrate

con1_plot=ggplot(d_pair6_con1, aes(x=time, y=HR1S)) +
  geom_line() +
  geom_line(aes(x=time, y=HR2S), colour="red")
con1_plot

con2_plot=ggplot(d_pair6_con2, aes(x=time, y=HR1S)) +
  geom_line() +
  geom_line(aes(x=time, y=HR2S), colour="red")
con2_plot

con3_plot=ggplot(d_pair6_con3, aes(x=time, y=HR1S)) +
  geom_line() +
  geom_line(aes(x=time, y=HR2S), colour="red")
con3_plot

gridExtra::grid.arrange(con1_plot, con2_plot, con3_plot, ncol=3)

#breathing

con1_B_plot=ggplot(d_pair6_con1, aes(x=time, y=Resp1S)) +
  geom_line() +
  geom_line(aes(x=time, y=Resp2S), colour="red")
con1_plot

con2_B_plot=ggplot(d_pair6_con2, aes(x=time, y=Resp1S)) +
  geom_line() +
  geom_line(aes(x=time, y=Resp2S), colour="red")
con2_plot

con3_B_plot=ggplot(d_pair6_con3, aes(x=time, y=Resp1S)) +
  geom_line() +
  geom_line(aes(x=time, y=Resp2S), colour="red")
con3_plot

gridExtra::grid.arrange(con1_B_plot, con2_B_plot, con3_B_plot, ncol=3)

```

  
- Can you eye-ball which condition if any displays more physiological coordination?

There's no coordination for heartrate, condition "syncronous", or the data has problems.

- Run crqa on heart rate and respiration data (find parameters, run crqa)

Heart rate
```{r}
library(crqa)
#Finding the optimal parameters: 
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 

ans= optimizeParam(d_pair6_con1$HR1S, d_pair6_con1$HR2S, par, min.rec= 3.5, max.rec= 4.5)

#Choose a common value for delay, emb.dim and radius
Results=crqa (d_pair6_con1$HR1S, d_pair6_con1$HR2S, delay=45, embed=3, radius=1, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)

#Represent the plot:
RP=Results$RP
RP = matrix(as.numeric(RP), nrow = ncol(RP))
cols = c("white","blue4")
image(RP, xlab = "", ylab = "", col = cols)

```

Breathing
```{r}
library(crqa)
#Finding the optimal parameters: 
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 

ans= optimizeParam(d_pair6_con1$Resp1S, d_pair6_con1$Resp2S, par, min.rec= 3.5, max.rec= 4.5)

#Choose a common value for delay, emb.dim and radius
Results=crqa (d_pair6_con1$Resp1S, d_pair6_con1$Resp2S, delay=45, embed=3, radius=1, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
#Represent the plot:
RP=Results$RP
RP = matrix(as.numeric(RP), nrow = ncol(RP))
cols = c("white","blue4")
image(RP, xlab = "", ylab = "", col = cols)

```

- Does this tell you more than just eyeballing the plots?

### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.

```{r functions}
library(groupdata2)
library(tidyverse)
library(ggplot2)

rm_outliers= function(alma){
  
  #remove outliers

  removeOuts <- function(ts,threshold){
    ts[ts > (mean(ts,na.rm=T) +
            (threshold*sd(ts,na.rm=T))) |
       ts < (mean(ts,na.rm=T) -
            (threshold*sd(ts,na.rm=T)))] =
    mean(ts,na.rm=T)   
    return(ts)
    }
  
  threshold=1.5
  
  alma$HR1=removeOuts(alma$HR1,threshold)
  alma$HR2=removeOuts(alma$HR2,threshold)
  alma$Resp1=removeOuts(alma$Resp1,threshold)
  alma$Resp2=removeOuts(alma$Resp2,threshold)
  
  
  return(alma)
  
}

downsample= function(alma){
  
  #downsampling

  alma = alma %>%    
    group(n= 100, method= 'greedy') %>%    
    dplyr::summarise(
      time= mean(time,na.rm=T),
      HR1 = mean(HR1,na.rm=T),
      HR2 = mean(HR2,na.rm=T),
      Resp1 = mean(Resp1,na.rm=T),
      Resp2 = mean(Resp2,na.rm=T))
  
  return(alma)
}  

scaling= function(alma){
  #scaling
  
  alma$Resp1S=scale(alma$Resp1)
  alma$Resp2S=scale(alma$Resp2)
  alma$HR1S=scale(alma$HR1) 
  alma$HR2S=scale(alma$HR2)
  
  return(alma)
}

# see=rm_outliers(pair6_con1)
# see=downsample(see)
# see=scaling(see)

#create function that has all the little ones inside to preprocess the data
preprocess=function(data) {
  final=rm_outliers(data) 
  final= downsample(final)
  final= scaling(final)
  return(final)
}


#get the info from filenames

#save the original code for reference
#library(magrittr)
# string="cleandata/Study2_G6_T1_TurnTaking.csv"
# pattern="[[:punct:]]" #regex for punctuation so I'll split the string by underscores
# name= string %>% stringr::str_split(pattern) %>% unlist()
# name[5]


get_name= function(string) {
  
  pattern="[[:punct:]]" #regex for punctuation so I'll split the string by underscores
  name= unlist(stringr::str_split(string, pattern))
  study= name[2]
  group_1=name[3]
  trial=name[4]
  condition= name[5]

  basic= data.frame(study, group, trial, condition) #save them to a dataframe
  return(basic)
}

#naem=get_name("cleandata/Study2_G6_T1_TurnTaking.csv") #works

get_name_sur= function(string) {
  
  pattern="[[:punct:]]" #regex for punctuation so I'll split the string by underscores
  name= unlist(stringr::str_split(string, pattern))
  study= name[2]
  group_1=unlist(stringr::str_extract_all(name[3], "\\d+"))
  group_2=unlist(stringr::str_extract_all(name[4], "\\d+"))
  combination= unlist(stringr::str_extract_all(name[4], "[a-z]"))
  condition= name[5]

  basic= data.frame(study, group_1, group_2, combination, condition) #save them to a dataframe
  return(basic)
}


```



```{r plot function}
library(ggplot2)

plotting_HR = function (data) {
  #plot HRs
  HR = ggplot(data, aes(x = time)) +
    geom_line(aes(y = HR1S), colour = "red") +
    geom_line(aes(y = HR2S)) 
  
  return(HR)
}

plotting_Resp = function(data) {
  #plot Respirations
  Resp=ggplot(data,aes(x=time))+
    geom_line(aes(y=Resp1S),colour="red")+
    geom_line(aes(y=Resp2S))
  
  return(Resp)
}

#plotting_HR(see)
#plotting_Resp(see)


file_list = list.files(path= "new_cleandata", full.names=T)
for (x in file_list){
  y=read.csv(x)
  
  less= substring(x, 14, 50)
  ido= unlist(regmatches(less, regexec("/(.+).csv", less)))
  nombre= ido[2]
  heart=plotting_HR(y)
  nombre=paste("plot_heart", nombre, ".png", sep= "_") 
  ggsave(nombre, plot=heart, device = "png")
  
  lessy= substring(x, 14, 50)
  odi= unlist(regmatches(lessy, regexec("/(.+).csv", lessy)))
  nev= odi[2]
  resp=plotting_Resp(y)
  nev=paste("plot_resp", nev, ".png", sep= "_")
  ggsave(nev, plot=resp, device = "png")
  
}

# basic="new_cleandata/new_Study2_G9_T10_apple.csv"
# nombre= substring(basic, 14, 50)
# ido= unlist(regmatches(nombre, regexec("/(.+).csv", nombre)))
# nombre= ido[2]

```


```{r create new_data}

file_list = list.files(path= "cleandata", full.names=T) # this is a list with the names of the files

store=data.frame() #create empty dataframe to store data


#loop through all the files
for (i in file_list){
  basic=get_name(i) #get out the info from the name
  nuevo= "new"
  nombre= substring(i, 11, 50) #save the name of the file
  nuevo_nombre=paste(nuevo, nombre, sep= "_") #save the name of the file with "new" written before it
  duck=read.csv(i)  
  done=preprocess(duck) 
  whole= cbind(basic, done) 
  write.csv(whole, file = nuevo_nombre, row.names = F) #write to csv to create the new file
  whole=data.frame() #empty the dataframe to use it again
}

```


- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.

```{r optimizeParam functions}

library(crqa)


cerka_heart=function(alma) {
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 
  
  ans= optimizeParam(alma$HR1S, alma$HR2S, par, min.rec= 3.5, max.rec= 4.5)
  
  if (is.null(ans)) {          
    ans=haz_empty()                
  }
  
  return(ans)
}

cerka_resp=function(alma) {
  par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip") 
  
  ans= optimizeParam(alma$Resp1S, alma$Resp2S, par, min.rec= 3.5, max.rec= 4.5)
  
  if (is.null(ans)) {          
    ans=haz_empty()                
  }
  
  
  return(ans)
}

haz_empty=function(){
    delay=NA
    emddim=NA
    radius=NA
    empty= data.frame(delay, emddim, radius)
    return(empty)
  }


# stupid=read.csv("new_cleandata/new_Study2_G9_T3_TurnTaking.csv")
# 
# heart_par= cerka_heart(stupid)
# if (is.null(heart_par)) {
#   heart_par=haz_empty()
# }
# 
# resp_par= cerka_resp(stupid)
# if (is.null(resp_par)) {
#   resp_par=haz_empty()
#   }




```


```{r param loop} 

file_list = list.files(path= "new_cleandata", full.names=T) # this is a list with the names of the files

#heart rate
param_heart=data.frame()
param_resp= data.frame()

for (i in file_list){
  duck=read.csv(i)  
  hr=try(cerka_heart(duck), silent = T)
  param_heart= rbind(param_heart, hr) 
  re=try(cerka_resp(duck), silent = T) 
  param_resp= rbind(param_resp, re)
  
}

```


```{r param results}

param_heart=na.omit(param_heart)
param_resp=na.omit(param_resp)

param_heart$delay= as.numeric(param_heart$delay)
param_heart$emddim= as.numeric(param_heart$emddim)
param_heart$radius= as.numeric(param_heart$radius)

param_resp$delay= as.numeric(param_resp$delay)
param_resp$emddim= as.numeric(param_resp$emddim)
param_resp$radius= as.numeric(param_resp$radius)

param_heart=na.omit(param_heart)
param_resp=na.omit(param_resp)

delay_h= median(param_heart$delay)
emddim_h= median(param_heart$emddim)
radius_h= median(param_heart$radius)

delay_r= median(param_resp$delay)
emddim_r= median(param_resp$emddim)
radius_r= median(param_resp$radius)

#after trial and error, correction of the radious so RR is around 4
radius_h= radius_h * 1.4
radius_r= radius_r * 1.05

parameter_heart=c(delay_h, emddim_h, radius_h)
parameter_resp=c(delay_r, emddim_r, radius_r)

write.csv(parameter_heart, file= "param_heart.csv", row.names = F)
write.csv(parameter_resp, file= "param_resp.csv", row.names = F)

parameter_heart=read.csv("param_heart.csv")
parameter_resp=read.csv("param_resp.csv")

parameter_heart
parameter_resp 

# #little code for correction of RR
# heart_normal=read.csv("heart_normal.csv")
# resp_normal= read.csv("resp_normal.csv")
# entre=mean(na.omit(heart_normal$RR)) # 0.707 *1.35= 3.31 ; *1.4= 3.97
# kozte=mean(na.omit(resp_normal$RR)) # 3.747 *1.1 = 4.48 ; *1.05 = 4.11
# entre=sd(na.omit(heart_normal$RR)) # 0.87 #2.364 #2.6
# kozte=sd(na.omit(resp_normal$RR)) #0.54 #61 #0.57
# entre=range(na.omit(heart_normal$RR)) #0.053 3.556 #9.872 - 0.844
# kozte=range(na.omit(resp_normal$RR)) #3.050 5.126
# 
# 
# #little code for correction of RR
# e=read.csv("heart_shuffled.csv")
# l= read.csv("resp_shuffled.csv")
# entre=mean(na.omit(e$RR)) # 0.124 (original * 1.4)
# kozte=mean(na.omit(l$RR)) # 3.69 (original*1.05)

```

```{r crqa functions}

crqa_heart=function(alma) {
  adat=read.csv(alma)
  Results=crqa (adat$HR1S, adat$HR2S, delay=delay_h, embed=emddim_h, radius=radius_h, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  
  RR= Results$RR
  DET= Results$DET
  maxL=Results$maxL
  L=Results$L
  ENTR=Results$ENTR
  LAM=Results$LAM
  TT=Results$TT
  
  study= unique(adat$study)
  group=unique(adat$group)
  trial=unique(adat$trial)
  condition= unique(adat$condition)

  basic= data.frame(study, group, trial, condition) 
  
  crqa=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)
  
  final_heart=merge(basic, crqa)
  
  return(final_heart)

}

crqa_resp=function(alma) {
  adat=read.csv(alma)
  Results=crqa (adat$Resp1S, adat$Resp2S, delay=delay_r, embed=emddim_r, radius=radius_r, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  
  RR= Results$RR
  DET= Results$DET
  maxL=Results$maxL
  L=Results$L
  ENTR=Results$ENTR
  LAM=Results$LAM
  TT=Results$TT
  
  study= unique(adat$study)
  group=unique(adat$group)
  trial=unique(adat$trial)
  condition= unique(adat$condition)

  basic= data.frame(study, group, trial, condition) 
  
  crqa=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)
  
  final_resp=merge(basic, crqa)
  
  return(final_resp)

}

#test
# ghost=crqa_heart("new_cleandata/new_Study2_G9_T3_TurnTaking.csv")
# monster=crqa_resp("new_cleandata/new_Study2_G9_T3_TurnTaking.csv")

```

```{r crqa loop for normal}
file_list = list.files(path= "new_cleandata", full.names=T) 

#heart rate
heart_normal=data.frame()
resp_normal= data.frame()

for (i in file_list){
  hr=crqa_heart(i)
  heart_normal= rbind(heart_normal, hr)
  re=crqa_resp(i)
  resp_normal= rbind(resp_normal, re)
  
}

write.csv(heart_normal, file="heart_normal.csv", row.names = F)
write.csv(resp_normal, file="resp_normal.csv", row.names = F)


```

### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
 
 
```{r data for shuffled}

#function to create shuffles variables
shuffle= function(alma) {
  alma$HR1S= sample(alma$HR1S) 
  alma$HR2S= sample(alma$HR2S) 
  alma$Resp1S= sample(alma$Resp1S)
  alma$Resp2S= sample(alma$Resp2S)
  return(alma)
}

#loop to shuffle all files (create new, shuffled files)
file_list = list.files(path= "new_cleandata", full.names=T)

for (i in file_list){
  nombre= substring(i, 15, 50) 
  nuevo_nombre=paste("shuffled", nombre, sep= "_") 
  duck=read.csv(i)  
  done=shuffle(duck)
  write.csv(done, file = nuevo_nombre, row.names = F) 
}
```

```{r crqa loop for shuffled}
#loop for crqa of shuffled files
file_list = list.files(path= "shuffled_data", full.names=T)


heart_shuffled=data.frame()
resp_shuffled= data.frame()

for (i in file_list){
  hr=crqa_heart(i)
  heart_shuffled= rbind(heart_shuffled, hr)
  re=crqa_resp(i)
  resp_shuffled= rbind(resp_shuffled, re)

}

write.csv(heart_shuffled, file="heart_shuffled.csv", row.names = F)
write.csv(resp_shuffled, file="resp_shuffled.csv", row.names = F)

```


```{r}
#read in files
heart_normal=read.csv("heart_normal.csv")
resp_normal= read.csv("resp_normal.csv")
heart_shuffled= read.csv("heart_shuffled.csv")
resp_shuffled= read.csv("resp_shuffled.csv")

```


Statistically compare...

```{r}

heart_normal$type= factor("normal")
resp_normal$type= factor("normal")
heart_shuffled$type= factor("shuffled")
resp_shuffled$type= factor("shuffled")

#found this too late, might have been good
# When you supply a column name with the `.id` argument, a new
# column is created to link each row to its original data frame
#dplyr::bind_rows(list(one, two), .id = "id")
#bind_rows("group 1" = one, "group 2" = two, .id = "groups")

heart_normal_shuffled= rbind(heart_shuffled, heart_normal)
resp_normal_shuffled= rbind(resp_shuffled, resp_normal)

#relevel to have normal as baseline
heart_normal_shuffled$type= relevel(heart_normal_shuffled$type, "normal", "shuffled")
resp_normal_shuffled$type= relevel(resp_normal_shuffled$type, "normal", "shuffled")

library(lmerTest)

RR_h_lmer1=lmer(RR ~ type + (1|group), heart_normal_shuffled)
RR_r_lmer1=lmer(RR ~ type + (1|group), resp_normal_shuffled)

summary(RR_h_lmer1) #significant  -3.8562     0.4603 35.2500  -8.378 6.64e-10
summary(RR_r_lmer1) #significant -0.4195     0.1301 34.6800  -3.225  0.00275

DET_h_lmer1=lmer(DET ~ type + (1|group), heart_normal_shuffled)
DET_r_lmer1=lmer(DET ~ type + (1|group), resp_normal_shuffled)

summary(DET_h_lmer1) #-94.0336     0.7159  25.8090  -131.3  < 2e-16
summary(DET_r_lmer1) #-65.269      1.509  35.560  -43.27   <2e-16

LAM_h_lmer1=lmer(LAM ~ type + (1|group), heart_normal_shuffled)
LAM_r_lmer1=lmer(LAM ~ type + (1|group), resp_normal_shuffled)

summary(LAM_h_lmer1) #-95.9812     0.4436  35.5000  -216.4   <2e-16 
summary(LAM_r_lmer1) #-75.7001     1.1959  35.1900  -63.30   <2e-16

L_h_lmer1=lmer(L ~ type + (1|group), heart_normal_shuffled)
L_r_lmer1=lmer(L ~ type + (1|group), resp_normal_shuffled)

summary(L_h_lmer1) #-9.5718     0.6238 35.0600  -15.35  < 2e-16
summary(L_r_lmer1) # -1.28708    0.10168 35.23000  -12.66 1.15e-14

maxL_h_lmer1=lmer(maxL ~ type + (1|group), heart_normal_shuffled)
maxL_r_lmer1=lmer(maxL ~ type + (1|group), resp_normal_shuffled)

summary(maxL_h_lmer1) #-192.41      16.58   35.18  -11.61 1.40e-13
summary(maxL_r_lmer1) #-29.682      1.688  35.380  -17.58  < 2e-16

ENTR_h_lmer1=lmer(ENTR ~ type + (1|group), heart_normal_shuffled)
ENTR_r_lmer1=lmer(ENTR ~ type + (1|group), resp_normal_shuffled)

summary(ENTR_h_lmer1) #-2.42622    0.07389 25.51100  -32.83  < 2e-16
summary(ENTR_r_lmer1) #-1.35837    0.04916 35.36000  -27.63   <2e-16

TT_h_lmer1=lmer(TT ~ type + (1|group), heart_normal_shuffled)
TT_r_lmer1=lmer(TT ~ type + (1|group), resp_normal_shuffled)

summary(TT_h_lmer1) #-12.3219     0.8042  34.2200  -15.32  < 2e-16
summary(TT_r_lmer1) #-2.1697     0.1633 35.3500  -13.29 2.66e-15

#should be significant, except for RR

```


### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Celine will share a commented script
 
```{r surrogate pairs}

setwd("/Users/Viki/Documents/egyetem/exp.met/port4/assignment4")
prepro_filelist = list.files(path="new_cleandata",pattern = "Study1", full.names = T) # load the preprocessed files
library(stringr)

num=stringr::str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

df$V2=as.character(df$V2)
pairs=as.numeric(unique(df$V2)) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Synchronous","TurnTaking")){ #loop through a list of condintions #,"SelfPaced"
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study1_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist #study2_G
  f2=paste0('Study1_G',SurrogateList$b[i])
  
  file1=try(prepro_filelist[grepl(f1,prepro_filelist) & grepl(co, prepro_filelist)]) #insert conditions as well.
  file2=try(prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)])
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv") 
  file4=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study1_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write.csv(d3,file.path("surrogate",file3))
  write.csv(d4,file.path("surrogate",file4))
  write.csv(d5,file.path("surrogate",file5))
  write.csv(d6,file.path("surrogate",file6))
  }
}

##NOW DO THE SAME FOR STUDY 2##

setwd("/Users/Viki/Documents/egyetem/exp.met/port4/assignment4")
prepro_filelist = list.files(path="new_cleandata",pattern = "Study2", full.names = T) # load the preprocessed files

num=stringr::str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

df$V2=as.character(df$V2)
pairs=as.numeric(unique(df$V2)) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Synchronous","TurnTaking","Conversation")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('Study2_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('Study2_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1S,Resp2S=d2$Resp2S,HR1S=d1$HR1S,HR2S=d2$HR2S)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1S,Resp2S=d1$Resp2S,HR1S=d2$HR1S,HR2S=d1$HR2S)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1S,Resp2S=d2$Resp1S,HR1S=d1$HR1S,HR2S=d2$HR1S)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2S,Resp2S=d1$Resp2S,HR1S=d2$HR2S,HR2S=d1$HR2S)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv") 
  file4=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write.csv(d3,file.path("surrogate",file3))
  write.csv(d4,file.path("surrogate",file4))
  write.csv(d5,file.path("surrogate",file5))
  write.csv(d6,file.path("surrogate",file6))
  }
}
```

```{r s_crqa functions}

s_crqa_heart=function(alma) {
  adat=read.csv(alma)
  Results=crqa (adat$HR1S, adat$HR2S, delay=delay_h, embed=emddim_h, radius=radius_h, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  
  RR= Results$RR
  DET= Results$DET
  maxL=Results$maxL
  L=Results$L
  ENTR=Results$ENTR
  LAM=Results$LAM
  TT=Results$TT
  
  final_heart=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)
  
  return(final_heart)

}

s_crqa_resp=function(alma) {
  adat=read.csv(alma)
  Results=crqa (adat$Resp1S, adat$Resp2S, delay=delay_r, embed=emddim_r, radius=radius_r, normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
  
  RR= Results$RR
  DET= Results$DET
  maxL=Results$maxL
  L=Results$L
  ENTR=Results$ENTR
  LAM=Results$LAM
  TT=Results$TT
  
  final_resp=data.frame(RR, DET, maxL, L, ENTR, LAM, TT)
  
  
  return(final_resp)

}

#test
# ghost=s_crqa_heart("new_cleandata/new_Study2_G9_T3_TurnTaking.csv")
# monster=s_crqa_resp("new_cleandata/new_Study2_G9_T3_TurnTaking.csv")

```
 
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs
 
 
```{r surrogate loop}

#loop for crqa of surrogate files
file_list = list.files(path= "surrogate", full.names=T)


heart_surrogate=data.frame()
resp_surrogate= data.frame()

for (i in file_list){
  basic=get_name_sur(i)
  hr=s_crqa_heart(i)
  all=cbind(basic, hr)
  heart_surrogate= rbind(heart_surrogate, all)
  
  
  basicr=get_name_sur(i)
  re=s_crqa_resp(i)
  allr=cbind(basicr, re)
  resp_surrogate= rbind(resp_surrogate, allr)
  rm(all)
  rm(allr)
}

write.csv(heart_surrogate, file="heart_surrogate.csv", row.names = F)
write.csv(resp_surrogate, file="resp_surrogate.csv", row.names = F)

```


Statistically compare...

```{r}
heart_surrogate=read.csv("heart_surrogate.csv")
resp_surrogate=read.csv("resp_surrogate.csv")

heart_surrogate$type= factor("surrogate")
resp_surrogate$type= factor("surrogate")

heart_surrogate_mod=heart_surrogate
heart_surrogate_mod$group= paste(heart_surrogate_mod$group_1, heart_surrogate_mod$group_2)
heart_surrogate_mod$group=as.factor(heart_surrogate_mod$group)
heart_surrogate_mod= subset(heart_surrogate_mod, select= -c(group_1, group_2, combination))
heart_normal_mod= subset(heart_normal, select= -c(trial))

resp_surrogate_mod=resp_surrogate
resp_surrogate_mod$group= paste(resp_surrogate_mod$group_1, resp_surrogate_mod$group_2)
resp_surrogate_mod$group=as.factor(resp_surrogate_mod$group)
resp_surrogate_mod= subset(resp_surrogate_mod, select= -c(group_1, group_2, combination))
resp_normal_mod= subset(resp_normal, select= -c(trial))

heart_normal_surrogate= rbind(heart_surrogate_mod, heart_normal_mod)
resp_normal_surrogate= rbind(resp_surrogate_mod, resp_normal_mod)

#relevel to have normal as baseline
heart_normal_surrogate$type= relevel(heart_normal_surrogate$type, "normal", "surrogate")
resp_normal_surrogate$type= relevel(resp_normal_surrogate$type, "normal", "surrogate")

library(lmerTest)

RR_h_lmer=lmer(RR ~ type + (1|group), heart_normal_surrogate)
RR_r_lmer=lmer(RR ~ type + (1|group), resp_normal_surrogate)

summary(RR_h_lmer) #0.01269    0.77525 82.73000   0.016    0.987
summary(RR_r_lmer) #0.05708    0.11816 172.19000   0.483     0.63 

DET_h_lmer=lmer(DET ~ type + (1|group), heart_normal_surrogate)
DET_r_lmer=lmer(DET ~ type + (1|group), resp_normal_surrogate)

summary(DET_h_lmer) #-0.5071     0.9208  99.9800  -0.551    0.583
summary(DET_r_lmer)  #-0.07856    2.07370  90.50000  -0.038     0.97

LAM_h_lmer=lmer(LAM ~ type + (1|group), heart_normal_surrogate)
LAM_r_lmer=lmer(LAM ~ type + (1|group), resp_normal_surrogate)

summary(LAM_h_lmer) #-0.01492    0.60093 135.25000  -0.025     0.98
summary(LAM_r_lmer) #0.6963     1.3613 142.2900   0.512     0.61

L_h_lmer=lmer(L ~ type + (1|group), heart_normal_surrogate)
L_r_lmer=lmer(L ~ type + (1|group), resp_normal_surrogate)

summary(L_h_lmer) #0.2431     1.2304 129.3100   0.198    0.844 
summary(L_r_lmer) # -0.04365    0.11425 104.04000  -0.382    0.703

maxL_h_lmer=lmer(maxL ~ type + (1|group), heart_normal_surrogate)
maxL_r_lmer=lmer(maxL ~ type + (1|group), resp_normal_surrogate)

summary(maxL_h_lmer) # -6.111     22.151  71.370  -0.276    0.783
summary(maxL_r_lmer) #  0.1911     1.9475 117.7100   0.098    0.922  

ENTR_h_lmer=lmer(ENTR ~ type + (1|group), heart_normal_surrogate)
ENTR_r_lmer=lmer(ENTR ~ type + (1|group), resp_normal_surrogate)

summary(ENTR_h_lmer) #-0.02217    0.08912 102.94000  -0.249    0.804 
summary(ENTR_r_lmer) #-0.02013    0.05696 108.41000  -0.353    0.725

TT_h_lmer=lmer(TT ~ type + (1|group), heart_normal_surrogate)
TT_r_lmer=lmer(TT ~ type + (1|group), resp_normal_surrogate)

summary(TT_h_lmer) # 1.345      1.950 169.250   0.690    0.491
summary(TT_r_lmer) # 0.03624    0.19299 155.63000   0.188    0.851 

#no significant!
```


### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?
 
question 3: do interaction between condition and baseline
condition: sync, turn, conversation
type: real, shuffled, surrogated 
-> y ~ C*type 
 
```{r}

RR_h_int=lmer(RR ~ condition*type + (1|group), heart_normal_shuffled)
RR_r_int=lmer(RR ~ condition*type + (1|group), resp_normal_shuffled)

summary(RR_h_int)
#  Estimate Std. Error      df t value Pr(>|t|)
# conditionSynchronous:typeshuffled  -1.0483     1.1896 31.2300  -0.881 0.384959
# conditionTurnTaking:typeshuffled   -0.4290     1.1896 31.2300  -0.361 0.720821

summary(RR_r_int) 
# conditionSynchronous:typeshuffled -0.21982    0.28794 30.78000  -0.763    0.451    
# conditionTurnTaking:typeshuffled   0.05653    0.28794 30.78000   0.196    0.846  

DET_h_int=lmer(DET ~ condition*type + (1|group), heart_normal_shuffled)
DET_r_int=lmer(DET ~ condition*type + (1|group), resp_normal_shuffled)

summary(DET_h_int)
# conditionSynchronous:typeshuffled   0.9195     1.7066  19.0850   0.539    0.596    
# conditionTurnTaking:typeshuffled    1.4063     1.8188  19.9570   0.773    0.448  

summary(DET_r_int)
# conditionSynchronous:typeshuffled  -4.66522    3.79801  31.63000  -1.228   0.2284    
# conditionTurnTaking:typeshuffled   -0.69464    3.79801  31.63000  -0.183   0.8560  

LAM_h_int=lmer(LAM ~ condition*type + (1|group), heart_normal_shuffled)
LAM_r_int=lmer(LAM ~ condition*type + (1|group), resp_normal_shuffled)

summary(LAM_h_int)
# conditionSynchronous:typeshuffled   0.4199     1.1310  31.5800    0.371    0.713    
# conditionTurnTaking:typeshuffled    1.0347     1.1310  31.5800    0.915    0.367

summary(LAM_r_int)
# conditionSynchronous:typeshuffled   -1.288      3.119  31.200  -0.413    0.682    
# conditionTurnTaking:typeshuffled    -1.294      3.119  31.200  -0.415    0.681

L_h_int=lmer(L ~ condition*type + (1|group), heart_normal_shuffled)
L_r_int=lmer(L ~ condition*type + (1|group), resp_normal_shuffled)

summary(L_h_int)
# conditionSynchronous:typeshuffled -0.06999    1.65383 31.07000  -0.042    0.967    
# conditionTurnTaking:typeshuffled   0.66605    1.65383 31.07000   0.403    0.690

summary(L_r_int)
# conditionSynchronous:typeshuffled -0.56119    0.22142 31.37000  -2.535  0.01646 *  
# conditionTurnTaking:typeshuffled  -0.06176    0.22142 31.37000  -0.279  0.78213 

maxL_h_int=lmer(maxL ~ condition*type + (1|group), heart_normal_shuffled)
maxL_r_int=lmer(maxL ~ condition*type + (1|group), resp_normal_shuffled)

summary(maxL_h_int)
# conditionSynchronous:typeshuffled  -32.750     42.838   31.210  -0.765    0.450    
# conditionTurnTaking:typeshuffled     8.250     42.838   31.210   0.193    0.849

summary(maxL_r_int)
# conditionSynchronous:typeshuffled   -4.375      4.388  31.380  -0.997    0.326    
# conditionTurnTaking:typeshuffled    -1.625      4.388  31.380  -0.370    0.714 

ENTR_h_int=lmer(ENTR ~ condition*type + (1|group), heart_normal_shuffled)
ENTR_r_int=lmer(ENTR ~ condition*type + (1|group), resp_normal_shuffled)

summary(ENTR_h_int)
# conditionSynchronous:typeshuffled  0.07183    0.18131 19.10600   0.396    0.696    
# conditionTurnTaking:typeshuffled   0.09296    0.19333 19.93000   0.481    0.636

summary(ENTR_r_int)
# conditionSynchronous:typeshuffled -0.26581    0.10587 31.45000  -2.511  0.01740 *  
# conditionTurnTaking:typeshuffled  -0.03825    0.10587 31.45000  -0.361  0.72030 

TT_h_int=lmer(TT ~ condition*type + (1|group), heart_normal_shuffled)
TT_r_int=lmer(TT ~ condition*type + (1|group), resp_normal_shuffled)

summary(TT_h_int)
# conditionSynchronous:typeshuffled  -0.06821    1.99905  30.19000  -0.034    0.973    
# conditionTurnTaking:typeshuffled    1.77831    2.03347  30.31000   0.875    0.389

summary(TT_r_int)
# conditionSynchronous:typeshuffled  -0.7952     0.3729 31.1500  -2.133  0.04095 *  
# conditionTurnTaking:typeshuffled   -0.1382     0.3729 31.1500  -0.371  0.71337


```

### Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
 
 
 model1=heartrate~ condition
 model2=heartrate ~ condition + respiration
 anova(model1, model2)
 See whether adding respiration as a predictor made the model significantly better
 

 
 
 
 
 